---
title: "QRM II Graded Assignment (13)"
subtitle: "Period 1, 2026"
author: "Fill in your group number and names here"
date:  "`r format(Sys.Date(), '%d-%m-%Y')`"
output:
  pdf_document:
    number_sections: true
  html_document:
    df_print: paged

header-includes:
- "\\setcounter{section}{-1}"
- |
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
    showspaces = false,
    showtabs = false,
    breaklines,
    commandchars=\\\{\}
  }
format:
  html:
    code-overflow: wrap
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This assignment is to be completed in groups of 4-5. Further, all students in your group need to be assigned to the same R tutorial group (Friday's tutorial). You can sign yourself up for a group on Canvas. Please do so \textbf{before the start of your first R tutorial on Friday September 4th.} You can use the Discussion Board in Canvas if you do not have a group yet or if your group is incomplete.

The assignment has 5 parts, and each part corresponds to the course material of that week (with the exclusion of week 6, for which there is no R programming material). 

You are supposed to hand in these assignments on Canvas at the following dates:

  - **Deadline 1** *Sunday September 27th, at 23:59pm*: you are supposed to hand in weeks 1, 2, and 3 of this assignment.   This will determine 18\% of your overall course grade
  - **Deadline 2** *Sunday October 11th, at 23:59pm*: you are supposed to hand in weeks 4, and 5 of this assignment. This will determine 12\% of your overall course grade 

The R tutorials (each Friday) will consist of two halves. During the first half, you will discuss the tutorial exercises. These can be downloaded separately from Canvas. During the second half, you can work on this graded assignment within your own group. The purpose is that you find out how to work with R for doing statistical analyses by yourself. The tutorial exercises are meant to teach you basic commands to get you started, but to answer the problem sets in this assignment, you might need to research your own solutions, and use functions and commands not described in the tutorial exercises. Learning how to solve your own research problems is integral part of learning R. When you and your group get stuck on how to approach an exercise, the hierarchy in finding your way is as follows:

-	use the concepts from the tutorial exercises;
- use the cheat sheets available on Canvas;
-	use Google, YouTube, StackOverflow, or another website;
- ask the teacher.

The use of generative AI is \textbf{not} permitted and may result in a grade of 0. See the AI protocol in the course manual for details.

To answer the assignment, you can simply fill out this R markdown document. There are designated places which you can fill with R code. There are also designated spaces for you to answer each question. Often, the structure of an answer will be as follows. First, you type the R code in the designated box. This will show how you analyzed the data to get the answer to the question. Below the box for the R code, you will then summarize your answer to the question, i.e. what are the conclusions that you draw from the data analysis? 

When handing in, you are supposed to submit this .Rmd file, and a knitted version of this document. You can knit this document to pdf, word, or html. Knitting to pdf requires you to have a .tex distribution installed on your computer. Knitting to Word requires you to have Word installed. 

The exercises are designed such that you should be able to finish the majority of them during the tutorial each week. If you are not able to finish them fully during that time, you are expected to work on it in your own time using the computers on campus or your own device. It is best to meet as a group in-person when working together. If you want to work remotely, github is a good platform to guarantee smooth collaboration. Alternatively, you can email this .Rmd file back and forth to one another as a group, but this is not recommended as it is more cumbersome.

We encourage you to keep your code blocks, printing statements, and final answers, as short as possible. In any case, there is a page limit of 6 pages per week, which encompasses the total length of this document which consists of the questions, your coding lines, and your answers. When your answers to questions of the respective week exceed this page limit, they will not be graded, resulting in zero points.

Each week consists of 1, 2, or 3 subquestions. The total amount of points you can earn per week is 20 points. 

# Week 1 

1.  Find the dataset “movies1.tsv” on Canvas. Describe your data set: How many observations does it have. How many variables are there? How many subjects? What consists of a subject? \textbf{[4 points]}

```{r}
#WRITE YOUR CODE HERE
```

**Your Answer:**

Write your response here.

2. Which of the following types of variables are present in your data set? (i) nominal; (ii) ordinal; (iii); interval; (iv) ratio. If present, name one example of such a variable present in your data set.  \textbf{[4 points]} 

```{r}
#WRITE YOUR CODE HERE
head(movies1, n=5)
```

**Your Answer:**

rite your response here.


3. A movie studio wants to know which types of movies are best at generating profit. The measure they use for this is the profitability ratio, which is defined as $$PR=\frac{Revenue}{Budget}.$$ Perform the following steps to provide the movie studio with an analysis which corresponds to their request:

a. Create the variable PR as the revenue of a movie divided by its budget. Report its mean, median, maximum, and minimum.  \textbf{[2 points]}
b. As you can probably see from your answer in a.), some movies have a PR of "Inf". What does this mean? Why are some movies assigned this value? \textbf{[2 points]}
c. As you can probably see from your answer in a.), some movies have a PR of "NA". What does this mean? Why are some movies assigned this value? \textbf{[2 points]}
d. For the movies that have a PR of "Inf", set its PR to "NA". After doing this, report the mean, median, maximum, and minimum of PR. How do you interpret the mean? \textbf{[2 points]}
e. What does a PR of at least one mean? What is the fraction of movies that have a PR of at least one? \textbf{[2 points]}
f. Create a boxplot of the variable PR. Make sure it has an appropriate title, and appropriate titles and labels for the x- and y-axis. Describe the shape of the boxplot. Give $Q_1$, $Q_2$, $Q_3$ and $Q_4$. What does this tell you about the nature of making money in the movies industry? \textbf{[2 points]}
 
For each step, you should provide first all the code you used to answer the question and then formulate an answer using full sentences.

*Step a*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*Step b*

```{r}

#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*Step c*

```{r}

#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*Step d*

```{r}

#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*Step e*

```{r}

#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*Step f*

```{r}

#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

# Week 2 
1 Is your dataset movies1.tsv the full population, or is it a sample of a larger population? If the latter, how would you describe the full population? \textbf{[4 points]} 

**Your Answer**

Write your formulated response here.


2 Let M be the event that a movie director is male. Let R be the runtime of the movie. 

a. Using your data, calculate $P(M)$. What is $P(M')$?. Calculate $P(M')$ as well. \textbf{[1 point]} 
b. Assume R is normally distributed with $\mu=107$ and $\sigma=22$, calculate $P(90 \leq R \leq 120)$. \textbf{[1 point]} 
c. Calculate $P(90 \leq R \leq 120)$ empirically using your data. How accurate were the assumptions about the distribution of R in 2b? Which of the assumptions do you consider the most suspect? Explain. \textbf{[2 points]} 
d. Using your data, calculate $P(120 \leq R | M')$ \textbf{[2 points]} 

*step a*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step c*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step d*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

3 For this question, you will again use the profitability ratio PR. Make sure that PR is defined the same way as in question 3d of week 1. For this question, you will assume that your data set is the full population.  

a.	What is the mean of the variable PR in your data set?  \textbf{[2 points]} 
b.	Create a new data set, called movies_sample. Make sure that it is a random sample of your data set of 25 movies. What is the mean of PR in this random sample? How does it compare to the mean of PR in 3a?  \textbf{[2 points]} 
c.	In a for loop, create 100 different samples of 25 movies, as in b, and estimate the mean within each sample. Save the mean of each sample in a vector called sample_means. So the first position of the vector would have the mean of the first sample, the second position the mean of the second sample, etc. Print this vector.  \textbf{[2 points]} 
d.	Summarize and make a histogram of sample_means. What is the mean, standard deviation and shape of its distribution? \textbf{[2 points]} 
e. Using the formulas of the *Central Limit Theorem*, give the distribution of the sample mean of PR in a sample of 25 movies from you population (including its expected value and standard deviation). How do these values compare to your answers at d.)? \textbf{[2 points]} 

*step a*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step c*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step d*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step e*

**Your Answer:**

Write your formulated response here.

# Week 3 

1. For the next part of the assignment, assume that the movies in your data frame are a random sample of a larger population of movies. 

a. Create a new data set that only includes movies that are of the genre "Thriller". For these thriller movies, give a 99 percent confidence interval for the variable *vote_average*. \textbf{[2 points]} 

b. According to IMDB, the average movie is rated with a 6.5. Using your data, test for the null hypothesis that thriller movies are at least as good as the average movie, against the alternative hypothesis that thriller movies are rated worse than the average movie. Test this hypothesis using an approporiate five-step procedure. What is your p-value and what do you conclude? \textbf{[2 points]}

*step a*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

2

a. How many movies in your sample had a female lead actor? How many had a male lead actor? \textbf{[2 points]}
b. Test the null hypothesis that for any given movie the chance that the lead actor is male or female is equal. Clearly state your null hypothesis and alternative hypothesis, and your chosen significance level. Use an appropriate R function to test your hypothesis, and state your conclusion. Also give the p-value that corresponds to your test statistic. \textbf{[2 points]}
c. What is the variance of the profitability ratio (PR) for movies with a male lead actor? What is the variance of profitability ratio for movies with a female lead actor? Make sure that PR is defined in the same way as in week 1. \textbf{[2 points]}
d. Assume that the variance of PR for movies with a male lead actor in your data is the same as the variance for movies with a male lead actor in your population (that is, there is no sampling uncertainty in this estimate). Now, set up a hypothesis test to test for the null hypothesis that the PR for movies with a female lead is equal to the variance for movies with a male lead actor. Clearly state your null hypothesis and alternative hypothesis, and your chosen significance level. Calculate your test statistic and tell how your test statistic is distributed, and state your conclusion. Also give the p-value that corresponds to your test statistic. \textbf{[2 points]}


*step a*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.


*step c*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

*step d*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

3

a. Create a new variable called genre2. Make sure that the three most popular genres remain the same, but that all the other genres get categorized into one genre called "Other". Report a frequency table of this new variable genre2. \textbf{[2 points]} 
b. Create a scatter plot with each type of genre2 on the x axis and the mean of PR within that genre on the y-axis. Make sure it has an appropriate title, and appropriate titles and labels for the x- and y-axis. Which movie genre has the highest PR? Which has the lowest? \textbf{[2 points]}
c. Recreate the scatter plot, but now add bars around each point, indicating the 95\% confidence interval. \textbf{[2 points]}
d. Now, create a new data frame which is a random subsample of n=100 of your full data frame. Recreate the plot in c.) How do the confidence intervals compare to the confidence intervals you plotted in c.)? \textbf{[2 points]}


*step a*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.


*step c*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

*step d*

```{r}
#WRITE YOUR CODE HERE


```

**Your Answer:**

Write your formulated response here.

# Week 4 

1. There is an argument going on in the movie studio. *Bob* claims that production budgets are getting out of hand, and that the studio should focus on making cheaper movies. *Chantal* disagrees. She tells Bob that ``The profitability ratio of the movie will remain similar, no matter how much the budget is''. 

a. Set up a regression model to test Chantal's claim, and estimate it. That is, estimate: $$\text{PR}_i=\beta_0+\beta_1 \text{Budget}_i +\varepsilon_i.$$ Print a summary of your estimated model. \textbf{[2 points]}
b. What is the estimated value of $\beta_1$ how do you interpet it? \textbf{[2 points]}
c. Test for the null hypothesis that $\beta_1 = 0$. Report the p-value and state your conclusion. \textbf{[2 points]}
d. Who do you think is correct? Bob or Chantal? What would you advise the movie studio to do? \textbf{[2 points]}
e. Make a histogram of the residuals of your model. Based on the histogram, do you have any reason to doubt the validity of your model? Explain. \textbf{[2 points]}
f. Estimate hatvalues for each observation that you used to estimate your model. Do any hatvalues exceed the critical value? Based on this analysis, do you hvae any reason to doubt the validity of your model? \textbf{[2 points]}
g. Based on your analysis in f.), estimate your model again, but remove any outliers. Print a summary of your model and again test for the null hypothesis that $\beta_1 = 0$. What do you conclude? Based on this new model, would your advice to the movie studio in question d. change? \textbf{[2 points]}

*step a*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step c*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step d*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step e*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step f*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step g*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

2.

a. Make a scatterplot with vote_average on the y-axis, and runtime on the x-axis. Within the same scatterplot, draw a straight line which summarizes the association between the average vote and the runtime. \textbf{[3 points]}
b. Using the insights from the scatterplot, your linear model, and your data, find an example of a movie that vastly overperformed. That means, find a movie that had much higher average vote than expected, given its runtime. For this movie, report its vote_average, its predicted vote_average, and its residual. \textbf{[3 points]}

*step a*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

# Week 5
  
a. Create a plot of the mean PR by year of release. Include 95\% confidence intervals around each mean Do you see any indication of a time trend in the PR? \textbf{[4 points]}
b. Estimate an OLS model which has as dependent variable the PR of a movie, and as independent variable the log of budget, a dummy for each level that genre2 can take, and the year of release. Show a summary of the resulting model and interpret each coefficient. \textbf{[4 points]}

The movie studio that you work is releasing a new movie in May of 2026. It will be a Comedy movie with a budget of 40,000,0000. It will have a runtime of 160 minutes. 

c. Estimate a model that is able to predict the revenue of this movie. Give its predicted revenue and include a 95% prediction interval. \textbf{[6 points]}
d. Plot the residuals against various independent variables in the model. Do you find any evidence of heteroskedasticity? \textbf{[2 points]}
e. The movie is set in Germany. The director wants to make the movie with only german-speaking actors. The executive producer is strongly against this, as she is afraid that movies that are not english spoken will have a low profit ratio. Estimate a model that can test whether the language spoken impacts the PR of this specific movie. Use an appropriate hypothesis test to test whether the language. What would you advise the movie studio regarding the language to be spoken? \textbf{[6 points]}


*step a*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step b*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step c*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step d*

```{r}
#WRITE YOUR CODE HERE

```

**Your Answer:**

Write your formulated response here.

*step e*

```{r}
#WRITE YOUR CODE HERE

```


