---
title: "assignment13 week 3"
output: html_document
---

```{r}
library(tidyverse)
movies <- read_tsv("movies1.tsv")

```

Question 1A
```{r}
thriller_movies<- movies %>%
filter(genre== "Thriller")
  
view(thriller_movies)
view(movies)
t.test(thriller_movies$vote_average, conf.level = 0.99)
```
With a 99% confidence interval for the variable vote_average the mean lies between 5.644845 and 6.299771. 


Question 1B
```{r}
t.test(thriller_movies$vote_average,
       mu = 6.5,
       alternative = "less",
       conf.level = 0.99)
qt(0.01, df= 64)
```
Null hypothesis states that the true mean of x is bigger, or equal to 6.5. 
Alternative hypothesis states that the true mean is less than 6.5
$\alpha$ = 0.01,
df = 64, which is more than 30, so the sampling distribution of the mean of x is approximately normal. 
The critical t-value is -2.386037 and the calculated t-value is -4.2782.
The calculated t-value is smaller than the critical t-value, so the null hypothesis is rejected. 
the p-value is 0.0000321. 
Since the calculated T-value is smaller than the critical t-value, we reject the null hypothesis and conclude there is sufficient evidence to conclude that thriller movies are not at least as good as the average movie. 


Question 2a
```{r}
movies %>%
  count(first_actor_gender)
```
130 movies with a female lead actor, and 352 movies with a male lead actor. 23 movies with no information on the gender.


Question 2b
```{r}
movies_without_na <- movies%>%
filter(!is.na(first_actor_gender))
males_mvs <- sum(movies_without_na$first_actor_gender== "male")
total_gender <- nrow(movies_without_na)

prop.test(x=males_mvs, n=total_gender, p=0.5, alternative= "two.sided")
```

Null hypothesis states that for any given movie the chance for a male or female lead actor is equal, so if π = 0.5.
Alternative hypothesis states that the chance is not equal, with  π not being 0.5.
$\alpha$ = 0.05
There is sufficient evidence to conclude that the chance for a male or female lead actor is not equal. The p-value is virtually zero. 

Question 2c
```{r}
movies$PR <- movies$revenue / movies$budget
mean(movies$PR, na.rm = TRUE)
median(movies$PR, na.rm = TRUE)
max(movies$PR, na.rm = TRUE)
min(movies$PR, na.rm = TRUE)
movies_without_na <- movies%>%
filter(!is.na(first_actor_gender), budget> 0)
female_mvs <-  movies_without_na %>% filter(first_actor_gender=="female")
male_mvs <- movies_without_na %>% filter(first_actor_gender=="male")
var(male_mvs$PR, na.rm = TRUE)
var(female_mvs$PR, na.rm = TRUE)
``` 
the variance of PR with a male lead is 99.82828
the variance of pr with a female lead is 1987.742




